---
title: "ggplot Exercise 3"
author: "Amy Perfors (adapted from Danielle Navarro)"
date: "18/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We're going to continue our analysis is of the "Sydney beaches dataset, which contains info about the weather on a variety of Sydney beaches at different times.

```{r, dataload, error=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
data_file <- here("data", "sydneybeaches3.csv")
beaches <- read_csv(file = data_file)
```

The plot below plots the amount of bacteria found on the beaches by date.

```{r scatterplot, error=FALSE, warning=FALSE}
ggplot(
  data = beaches,
  mapping = aes(
    x = date,
    y = log10(enterococci)
  ) 
) + 
geom_point(size=2) +
theme_bw()
```

Add a blue line to this plot. Make sure the line goes *under* the points.

```{r lineplot, error=FALSE, warning=FALSE}

```

Make a boxplot of the bacteria (on the y axis) by the name of the season (on the x axis). For extra credit, make sure the seasons are in chronological order.

```{r boxplot, error=FALSE, warning=FALSE}

```

Add jittered points to your boxplot, and colour them according to the name of the season. Make them slightly see-through for extra credit.

```{r jitteredplot, error=FALSE, warning=FALSE}

```

Make a plot with four facets, one for each season, that shows a histogram of the temperature in each season.. 

```{r hists, error=FALSE, warning=FALSE}

```

Change those histograms to density plots (hint: look up the geom called geom_density). For extra credit, colour them by season_name. (Hint: use the fill parameter)

```{r densities, error=FALSE, warning=FALSE}

```
