---
title: "ggplot Exercise 1"
author: "Amy Perfors (adapted from Danielle Navarro)"
date: "18/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this DIY exercise, we have an R Markdown document with missing code. Our analysis is of th "Sydney beaches dataset, which contains info about the weather on a variety of Sydney beaches at different times.

```{r, dataload, error=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
data_file <- here("data", "sydneybeaches3.csv")
beaches <- read_csv(file = data_file)
```

The plot below is a scatterplot showning temperature on the x-axis and rainfall on the y-axis. First, run this section of code and see what happens:

```{r scatterplot, error=FALSE, warning=FALSE}
ggplot(
  beaches, 
  aes(x=temperature, 
      y=rainfall)) + 
geom_point()
```

Now construct a new plot that makes the points larger and colours them according to season. (Question: why does it do something different for the season and season_name variables?)

```{r tempplot, error=FALSE, warning=FALSE}

```

Next we need to add a title. Create a new plot that uses the labs() function to add an informative title to the plot

```{r titleadd, error=FALSE, warning=FALSE}

```

Finally, add the bw theme and facet by season as well.

```{r facetadd, error=FALSE, warning=FALSE}

```